---
title: "Chapter3"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(data.table)
library(survival)
library(DiagrammeR)
library(regmedint)

```

# Chapter 3

## 3.1 Mediation

### 3.1.1 Assumption

1. No hidden exposure-outcome confounding\n
2. No hidden mediator-outcome confounding\n
3. No hidden exposure-mediator confounding\n	
4. No mediator–outcome confounding that is affected by the exposure\n

### 3.1.2 Basic estimation formula in parametric form

$$E (M|A = a,C = c) = \beta_0+\beta_1 a+ \beta_2^{'} c$$

$$E (Y |M = m,A = a,C = c) = \theta_0+\theta_1 a+ \theta_2 m + \theta_3 a m + \theta_4^{'} c$$

$$NDE = (\theta_1 + \theta_3 \beta_0 + \theta_3 \beta_1 a^* + \theta_3 \beta_2^{'} c)(a - a^*)$$
$$NIE = (\theta_2 \beta_1 + \theta_3 \beta_1 a)(a - a*)$$
### 3.1.3 Mediation with time-to-event outcome, AFT model

Addition assumption, mediator happened before outcome.

AFT model are able to handle both rare and common outcome

$$M =  \beta_0+\beta_1 a+ \beta_2^{'} c + \epsilon$$



$$log(T) = \theta_0+\theta_1 a+ \theta_2 m + \theta_3 a m + \theta_4^{'} c + v \epsilon$$

$$NDE = exp((\theta_1 + \theta_3 (\beta_0 + \beta_1 a^* +  \theta_2^{'} c + \theta_2 \sigma^2))(a - a^*) + 0.5 \theta_3 ^2 \sigma^2 (a^2 - a^{*2}))$$
$$NIE = exp((\theta_2 \beta_1 + \theta_3 \beta_1 a)(a - a*))$$


## 3.2 Mediation with measurement error

When interaction term does not exist(le Cessie et al., 2012). Measurement in the mediator can be addressed using the classical measurement error.Let $\tilde M $ be the mediator value measured and M the the true value, and U be the error term with normal distribution centered around zero.

$$\tilde M = M + U, U \sim Normal(0, \sigma_u)$$

Let $\lambda$ be the proportion of the total variance in $\tilde M $ explained by M. 

$$ \lambda = \frac{\sigma_{m|c}^2}{\sigma_{m|c}^2 + \sigma_{u}^2}$$

Estimation of the mediator with invention reminds the same. And estimation of the outcome can be written as:

$$E(Y | A = a, \tilde M = \tilde m, C =c) = \theta_0+\theta_1 a+ \theta_2 \tilde m  + \theta_4^{'} c$$

Parameterized in a AFT model: 

$$log(T) = \tilde \theta_0+ \tilde \theta_1 a+ \tilde \theta_2 \tilde m + \tilde \theta_4^{'} c + v \epsilon$$


$$ \theta_1 = \tilde \theta_1 - \tilde \theta_2(\frac{1}{\lambda}) \beta_1$$

$$\theta_2 = \frac{\tilde \theta_2}{\lambda}$$

$$E(Y | A = a, \tilde M = \tilde m, C =c) = \theta_0+\theta_1 a+ \theta_2 \tilde m + \theta_3 a \tilde m + \theta_4^{'} c$$

$$log(T) = \tilde \theta_0+ \tilde \theta_1 a+ \tilde \theta_2 \tilde m + \tilde \theta_3 a \tilde m + \tilde \theta_4^{'} c + v \epsilon$$

Adjustment works but not when the interaction term is involved. 



## 3.3 Mediation with measurement error and truncation

### 3.3.1 Simulation framework

Simulated baseline selection (truncation)
```{r}
n<-10000
t<-c(rep(0,n/2),rep(1,n/2))
beta1<-10
x0<- rnorm(n,2,2)#uniform will cause difference
x<-x0+beta1*t
theta0<-2
theta1<-3
theta2<-4
y<-theta0+theta1*t+theta2*x+rnorm(n,0,1)
w1<-x0+rnorm(n,0,1)
w2<-x+rnorm(n,0,1)

d1<-data.frame(y,t,w1,x,x0,label = "truth")
d2<-data.frame(y,t,w1,x = w2,x0 = x,label = "measured_b")
d<-rbind(d1,d2)

ggplot(data = d)+
  geom_point(aes(x,y,col = label),alpha = 0.2)+
  geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
  facet_wrap(~t)


d.sub<-d%>%filter(w1<2)%>%filter(label=="measured_b")
d.sub$label<-"measured.w/.cutoff"

d<-rbind(d,d.sub)
ggplot(data = d)+
  geom_point(aes(x,y,col = label),alpha = 0.2)+
  geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
  facet_wrap(~t)+
  ggtitle("")

d0.sub<-d.sub%>%filter(t==0)

df<-rbind(data.frame(x = d0.sub$w1,label = "baseline"),
          data.frame(x = d0.sub$x,label = "follow-up"),
          data.frame(x = d0.sub$x0,label = "truth"))
df$label<-as.factor(df$label)

# observed regression toward the mean
ggplot(data = df)+
  geom_histogram(aes(x = x,color = label),fill = "white",bins = 100)+
  ggtitle("Baseline vs follow-up vs trugh with truncation on baseline")


# hist(d0.sub$w1)
# hist(d0.sub$x)
# hist(d0.sub$x0)
# var(d0.sub$x)# should be 2
# var(d0.sub$x0)# use this variance in selected 

```


### 3.3.2 Biase and adjustment

```{r}
true.slope<-NULL
true.intercept<-NULL
true.t<-NULL

m.slope<-NULL
m.intercept<-NULL
m.t<-NULL

ms.slope<-NULL
ms.intercept<-NULL
ms.t<-NULL

sigma2.xs<-NULL
lambda.est<-NULL

ms.beta1<-NULL


for(i in 1:1000){
  n<-1000
  t<-c(rep(0,n/2),rep(1,n/2))
  beta1<-10
  x0<- rnorm(n,2,2)#uniform will cause difference
  x<-x0+beta1*t
  theta0<-2
  theta1<-3
  theta2<-4
  y<-theta0+theta1*t+theta2*x+rnorm(n,0,1)
  w1<-x0+rnorm(n,0,1)
  w2<-x+rnorm(n,0,1)
  
  d1<-data.frame(y,t,w1,x,x0,label = "truth")
  d2<-data.frame(y,t,w1,x = w2,x0 = x,label = "measured_b")
  d<-rbind(d1,d2)
  
  # ggplot(data = d)+
  #   geom_point(aes(x,y,col = label),alpha = 0.2)+
  #   geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
  #   facet_wrap(~t)
  # 
  
  d.sub<-d%>%filter(w1<2)%>%filter(label=="measured_b")
  d.sub$label<-"measured.w/.cutoff"
  
  d<-rbind(d,d.sub)
  # ggplot(data = d)+
  #   geom_point(aes(x,y,col = label),alpha = 0.2)+
  #   geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
  #   facet_wrap(~t)
  
  d0.sub<-d.sub%>%filter(t==0)


  
  sigma2.x<-var(d0.sub$x0)
  sigma2.xs[i]<-sigma2.x
  sigma2.u<-1
  
  
  lambda.est[i]<-sigma2.x / (sigma2.u + sigma2.x)
  
  
  
  ## With numbers, cut off play a role
  fit.m<-lm(x~t,data = d%>%filter(label == "measured.w/.cutoff"))
  
  fit1<-lm(y~x+t,data = d%>%filter(label == "truth"))
  fit2<-lm(y~x+t,data = d%>%filter(label == "measured_b"))
  fit3<-lm(y~x+t,data = d%>%filter(label == "measured.w/.cutoff"))
  
  ms.beta1[i]<-fit.m$coefficients[2]
  
  true.intercept[i]<-fit1$coefficients[1]
  true.slope[i]<-fit1$coefficients[2]
  true.t[i]<-fit1$coefficients[3]
  
  m.intercept[i]<-fit2$coefficients[1]
  m.slope[i]<-fit2$coefficients[2]
  m.t[i]<-fit2$coefficients[3]
  
  
  ms.intercept[i]<-fit3$coefficients[1]
  ms.slope[i]<-fit3$coefficients[2]
  ms.t[i]<-fit3$coefficients[3]
  
  
}

sigma2.x<-4
sigma2.u<-1
lambda<-sigma2.x / (sigma2.u + sigma2.x) #0.8
lambda.m<-mean(lambda.est)#~0.6611
beta1.m<-mean(ms.beta1)
# lambda.m
# 4*0.8
# 4*0.6611

# summary(true.slope)
# summary(m.slope)
# summary(ms.slope)
# 
# summary(m.intercept)
# summary(ms.intercept)
# 
# summary(true.t)
# summary(m.t)
# summary(ms.t)

df<-rbind(data.frame(value = true.slope,label = "truth"),
          # data.frame(value = m.slope/lambda,label = "adjusted.measured"),
          # data.frame(value = m.slope,label = "measured "),
          data.frame(value = ms.slope,label = "truncted measurements"),
          data.frame(value = ms.slope/lambda,label = "adjusted truncted measurements"),
          data.frame(value = ms.slope/lambda.m,label = "new adjusted truncted measurements"))

ggplot(data = df)+
  geom_histogram(aes(x = value,color = label),fill = "white",bins = 100)+
  ggtitle("New adjustment vs original on theta2")



df<-rbind(data.frame(value = true.t,label = "truth"),
          # data.frame(value = m.slope/lambda,label = "adjusted.measured"),
          # data.frame(value = m.slope,label = "measured "),
          data.frame(value = ms.t,label = "truncted measurements"),
          data.frame(value = (ms.t-ms.slope*(1/lambda-1)*beta1.m),label = "adjusted truncted measurements"),
          data.frame(value = (ms.t-ms.slope*(1/lambda.m-1)*beta1.m),label = "new adjusted truncted measurements"))

ggplot(data = df)+
  geom_histogram(aes(x = value,color = label),fill = "white",bins = 100)+
  ggtitle("New adjustment vs original on theta1")



```

## 3.4 Covariates and interaction term

### 3.5 Covariates dependent with mediator

In the previous formula, we know that when covariate c and m are not independent then

$$\sigma^2_{m}  \ge \sigma^2_{m|c}$$

$$ \lambda = \frac{\sigma_{m|c}^2}{\sigma_{m|c}^2 + \sigma_{u}^2} \le \frac{\sigma_{m}^2}{\sigma_{m}^2 + \sigma_{u}^2} = \lambda^*$$

Therefore, in cases that we are not able estimate $\sigma^2_{m|c} $ directly, we can still conclude an upper bound of the adjustment coefficient $\lambda$ using $\sigma^2_m$. Additionally, we can calculate the direction of reminding bias in $\theta_1$ and $\theta_2$

### 3.6 Exposure-mediator interaction

```{r}
rd32<-fread("v3-2.csv")
rd33<-fread("v3-3.csv")

rd32$cutoff<-as.factor(rd32$cutoff)
levels(rd32$cutoff)[5]<-"No cutoff"
rd32$cutoff.n<-as.numeric(rd32$cutoff)

rd33$cutoff<-as.factor(rd33$cutoff)
levels(rd33$cutoff)[5]<-"No cutoff"
rd33$cutoff.n<-as.numeric(rd33$cutoff)
```

#### theta1
```{r}

ggplot(data = rd32, aes(x = e1.theta1, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")




ggplot(data = rd33, aes(x = e1.theta1, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")
```

#### theta2
```{r}

ggplot(data = rd32, aes(x = e1.theta2, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")




ggplot(data = rd33, aes(x = e1.theta2, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")
```


####theta3
```{r}

ggplot(data = rd32, aes(x = e1.theta3, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")




ggplot(data = rd33, aes(x = e1.theta3, y = cutoff.n, fill = cutoff)) +
   geom_density_ridges(alpha = 0.5)+
 
  scale_y_discrete(limit = c(1:5),labels = data$cutoff)+
  coord_flip()+
  theme(legend.position = "bottom",axis.text.x = element_blank(),
        axis.title.x=element_blank())+
  xlab("Treatment on Mediator")+
  ggtitle("theta3")
```


```


## 4.

Let $\lambda^*$ be the upper limit of $\lambda$,which always smaller than 1, recall 

$$ \theta_1 = \tilde \theta_1 - \tilde \theta_2(\frac{1}{\lambda}) \beta_1$$

$$ \theta_1 \le \theta_1^* = \tilde \theta_1 - \tilde \theta_2(\frac{1}{\lambda^*}) \beta_1$$
Also recall

$$\theta_2 = \frac{\tilde \theta_2}{\lambda}$$

$$\theta_2 \ge\theta_2^* = \frac{\tilde \theta_2}{\lambda^*}$$ 

#


