---
title: "Chapter2"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)

```

# Appendix for Chapter Two

## 2.1 Regression toward the mean (RTM)

```{r}
#

x<- rnorm(1000,2,2)
w1<- x + rnorm(1000,0,1)
w2<- x + rnorm(1000,0,1)
#
d<-data.frame(x,w1,w2)
# selection
ds<-d%>%filter(w1<2)

summary(ds$w1)
summary(ds$w2)

t.test(ds$w1,ds$w2)

wilcox.test(ds$w1,ds$w2)


```

## 2.2 Classical measurment error in regression model

In classical measurement error framework, we assume X is the true value, U is the random error centered at 0, and W is observed measurement. 

$$W = X + U, U \sim Normal(0, \sigma_u)$$
In this scenario, the objective of regression can be written as

$$Y = \beta_0 + \beta_x  X + \beta_{z}^tZ + \epsilon$$

Coefficients $\beta_x$ can be estimated using regression function

$$Y = \beta_0^* + \beta_{w}  W + \beta_{z}^tZ + \epsilon$$


$$ \lambda = \frac{\sigma_{x|z}^2}{\sigma_{x|z}^2 + \sigma_{u}^2}$$

$$\beta_x = \beta_{w}^ \lambda$$
```{r}

## Classical measurement error

n<-1000
x<- rnorm(n,2,2)#uniform will cause more difference
#x<- runif(n,0,4)#uniform will cause more difference
beta0<-2
beta1<-4
y<-beta0+beta1*x+rnorm(n,0,1)
w<-x+rnorm(n,0,1)#w and y bi-variate normal

d1<-data.frame(y,x,label = "truth")
d2<-data.frame(y,x = w,label = "measured")
d<-rbind(d1,d2)

ggplot(data = d)+
  geom_point(aes(x,y,col = label),alpha = 0.2)+
  geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")

```




## 2.3 Classical measurment error in survival analysis

This framework can be extend to survival analysis. Recall common survival function:



$$ S(t) = exp(-\lambda t)$$

Solving for t gives:

$$ t = -log(S(t)) \frac{1}{\lambda}$$

Reparameterizing 

$$\frac{1}{\lambda} = exp(\beta^t X)$$

In a general form of AFT model, where x the variation has measurement error

$$log(T) = \beta_0 + \beta_x * X + \beta_{z}^t*Z + \epsilon$$




$$W = X + U, U \sim Normal(0, \sigma_u)$$

$$$$



## 2.4 Measurement error with truncation

### X is normally distributed

```{r}
set.seed(1)
n<-1000
x<- rnorm(n,2,2)#uniform will cause more difference
#x<- runif(n,0,4)#uniform will cause more difference
beta0<-2
beta1<-4
y<-beta0+beta1*x+rnorm(n,0,1)
w<-x+rnorm(n,0,1)#w and y bi-variate normal

d1<-data.frame(y,x,label = "truth")
d2<-data.frame(y,x = w,label = "measured")
d<-rbind(d1,d2)



d.sub<-d%>%filter(x<2)%>%filter(label=="measured")
d.sub$label<-"measured.w/.cutoff"

d<-rbind(d,d.sub)
ggplot(data = d)+
  geom_point(aes(x,y,col = label),alpha = 0.2)+
  geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
```




### Adjustment

#### Normal

```{r}
true.slope<-NULL
true.intercept<-NULL

m.slope<-NULL
m.intercept<-NULL

ms.slope<-NULL
ms.intercept<-NULL

for(i in 1:1000){
  #print(i)
  n<-1000
  alpha<-1
  t<-c(rep(0,n/2),rep(1,n/2))
  x<- rnorm(n,2,2) + alpha*t #uniform will cause difference
  beta0<-2
  beta1<-4
  y<-beta0+beta1*x+rnorm(n,0,1)
  w<-x+rnorm(n,0,1)#w and y bi-variate normal
  
  d1<-data.frame(y,x,label = "truth")
  d2<-data.frame(y,x = w,label = "measured")
  d<-rbind(d1,d2)
  
  ggplot(data = d)+
    geom_point(aes(x,y,col = label),alpha = 0.2)+
    geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
    ggtitle("X normally distributed")
  
  
  d.sub<-d%>%filter(x<2)%>%filter(label=="measured")
  d.sub$label<-"measured.w/.cutoff"
  
  d<-rbind(d,d.sub)
  # ggplot(data = d)+
  #   geom_point(aes(x,y,col = label),alpha = 0.2)+
  #   geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
  
  
  
  ## With numbers
  fit1<-lm(y~x,data = d%>%filter(label == "truth"))
  fit2<-lm(y~x,data = d%>%filter(label == "measured"))
  fit3<-lm(y~x,data = d%>%filter(label == "measured.w/.cutoff"))
  
  
  true.intercept[i]<-fit1$coefficients[1]
  true.slope[i]<-fit1$coefficients[2]
  
  m.intercept[i]<-fit2$coefficients[1]
  m.slope[i]<-fit2$coefficients[2]
  
  ms.intercept[i]<-fit3$coefficients[1]
  ms.slope[i]<-fit3$coefficients[2]
  
  
}



summary(true.slope)
summary(m.slope)
summary(ms.slope)

sigma2.x<-4
sigma2.u<-1


lambda<-sigma2.x / (sigma2.u + sigma2.x) #0.8
#lambda1<-sd(d.sub$x)^2 / (sigma2.u + sd(d.sub$x)^2) #0.8

df<-rbind(data.frame(value = true.slope,label = "truth"),
          data.frame(value = m.slope/lambda,label = "adjusted.measured"),
          data.frame(value = m.slope,label = "measured "),
          data.frame(value = ms.slope,label = "truncted measurements"),
          data.frame(value = ms.slope/lambda,label = "adjusted truncted measurements"))

ggplot(data = df)+
  geom_histogram(aes(x = value,color = label),fill = "white",bins = 100)+
  ggtitle("X normal, indepedent from Z")



```


### 2.5 X is not normally distributed

### X is uniformly distributed

#### Uniform

```{r}
set.seed(1)
n<-1000
#x<- rnorm(n,2,2)#uniform will cause more difference
x<- runif(n,0,4)#uniform will cause more difference
beta0<-2
beta1<-4
y<-beta0+beta1*x+rnorm(n,0,1)
w<-x+rnorm(n,0,1)#w and y bi-variate normal

d1<-data.frame(y,x,label = "truth")
d2<-data.frame(y,x = w,label = "measured")
d<-rbind(d1,d2)

d.sub<-d%>%filter(x<2)%>%filter(label=="measured")
d.sub$label<-"measured.w/.cutoff"

d<-rbind(d,d.sub)
ggplot(data = d)+
  geom_point(aes(x,y,col = label),alpha = 0.2)+
  geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")+
  ggtitle("X uniformlly distributed")
```


#### Without covariates

```{r}
true.slope<-NULL
true.intercept<-NULL

m.slope<-NULL
m.intercept<-NULL

ms.slope<-NULL
ms.intercept<-NULL

for(i in 1:1000){
  #print(i)
  n<-1000
  alpha<-1
  t<-c(rep(0,n/2),rep(1,n/2))
  x<- runif(n,0,4)  #uniform will cause difference
  beta0<-2
  beta1<-4
  y<-beta0+beta1*x+rnorm(n,0,1)
  w<-x+rnorm(n,0,1)#w and y bi-variate normal
  
  d1<-data.frame(y,x,label = "truth")
  d2<-data.frame(y,x = w,label = "measured")
  d<-rbind(d1,d2)
  
  ggplot(data = d)+
    geom_point(aes(x,y,col = label),alpha = 0.2)+
    geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
  
  
  d.sub<-d%>%filter(x<2)%>%filter(label=="measured")
  d.sub$label<-"measured.w/.cutoff"
  
  d<-rbind(d,d.sub)
  # ggplot(data = d)+
  #   geom_point(aes(x,y,col = label),alpha = 0.2)+
  #   geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
  
  
  
  ## With numbers
  fit1<-lm(y~x,data = d%>%filter(label == "truth"))
  fit2<-lm(y~x,data = d%>%filter(label == "measured"))
  fit3<-lm(y~x,data = d%>%filter(label == "measured.w/.cutoff"))
  
  
  true.intercept[i]<-fit1$coefficients[1]
  true.slope[i]<-fit1$coefficients[2]
  
  m.intercept[i]<-fit2$coefficients[1]
  m.slope[i]<-fit2$coefficients[2]
  
  ms.intercept[i]<-fit3$coefficients[1]
  ms.slope[i]<-fit3$coefficients[2]
  
  
}



summary(true.slope)
summary(m.slope)
summary(ms.slope)

sigma2.x<-4^2/12
sigma2.u<-1


lambda<-sigma2.x / (sigma2.u + sigma2.x) #0.8
#lambda1<-sd(d.sub$x)^2 / (sigma2.u + sd(d.sub$x)^2) #0.8

df<-rbind(data.frame(value = true.slope,label = "truth"),
          data.frame(value = m.slope/lambda,label = "adjusted.measured"),
          data.frame(value = m.slope,label = "measured "),
          data.frame(value = ms.slope,label = "truncted measurements"),
          data.frame(value = ms.slope/lambda,label = "adjusted truncted measurements"))

ggplot(data = df)+
  geom_histogram(aes(x = value,color = label),fill = "white",bins = 100)+
  ggtitle("X uniform, no covariates")



```


#### With covariates (independent)

```{r}
true.slope<-NULL
true.intercept<-NULL

m.slope<-NULL
m.intercept<-NULL

ms.slope<-NULL
ms.intercept<-NULL

for(i in 1:1000){
  #print(i)
  n<-1000
  alpha<-1
  t<-c(rep(0,n/2),rep(1,n/2))
  x<- runif(n,0,4) + t*alpha #uniform will cause difference
  beta0<-2
  beta1<-4
  y<-beta0+beta1*x+rnorm(n,0,1)
  w<-x+rnorm(n,0,1)#w and y bi-variate normal
  
  d1<-data.frame(y,x,label = "truth")
  d2<-data.frame(y,x = w,label = "measured")
  d<-rbind(d1,d2)
  
  ggplot(data = d)+
    geom_point(aes(x,y,col = label),alpha = 0.2)+
    geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
  
  
  d.sub<-d%>%filter(x<2)%>%filter(label=="measured")
  d.sub$label<-"measured.w/.cutoff"
  
  d<-rbind(d,d.sub)
  # ggplot(data = d)+
  #   geom_point(aes(x,y,col = label),alpha = 0.2)+
  #   geom_smooth(aes(x,y,col = label),method = "lm",formula = "y~x")
  
  
  
  ## With numbers
  fit1<-lm(y~x,data = d%>%filter(label == "truth"))
  fit2<-lm(y~x,data = d%>%filter(label == "measured"))
  fit3<-lm(y~x,data = d%>%filter(label == "measured.w/.cutoff"))
  
  
  true.intercept[i]<-fit1$coefficients[1]
  true.slope[i]<-fit1$coefficients[2]
  
  m.intercept[i]<-fit2$coefficients[1]
  m.slope[i]<-fit2$coefficients[2]
  
  ms.intercept[i]<-fit3$coefficients[1]
  ms.slope[i]<-fit3$coefficients[2]
  
  
}



summary(true.slope)
summary(m.slope)
summary(ms.slope)

sigma2.x<-4^2/12
sigma2.u<-1


lambda<-sigma2.x / (sigma2.u + sigma2.x) #0.8
#lambda1<-sd(d.sub$x)^2 / (sigma2.u + sd(d.sub$x)^2) #0.8

df<-rbind(data.frame(value = true.slope,label = "truth"),
          data.frame(value = m.slope/lambda,label = "adjusted.measured"),
          data.frame(value = m.slope,label = "measured "),
          data.frame(value = ms.slope,label = "truncted measurements"),
          data.frame(value = ms.slope/lambda,label = "adjusted truncted measurements"))

ggplot(data = df)+
  geom_histogram(aes(x = value,color = label),fill = "white",bins = 100)+
  ggtitle("X uniform, independent Z")



```

