---
title: "RGTM_Adjusting"
author: "Sicheng Hao"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(data.table)
library(survival)
library(DiagrammeR)
library(regmedint)

simulation.function.v42<-function(sample.size = 20000,cutoff = 1000){
  n<-sample.size/2
  osp = 2
  n0<-0
  n1<-0
  data0<-NULL
  data1<-NULL
  #control arm
  for(i in 1:100){
    #print(i)
    if(n0<n){
      ldlcm=rnorm(n = n*osp,mean = 116,sd = 14.2)#mean distribution of LDLC
      d0<-data.frame(ldlcm,ldlcb=rnorm(n = n*osp,mean = ldlcm,sd = 16))
      d0<-d0%>%filter(ldlcb<cutoff)
      data0<-rbind(data0,d0)
      n0<-nrow(data0)
    } else {
      break
    }
  }
  
  data0<-data0[1:n,]
  data0$ldlc12m<-data0$ldlcm+rnorm(n,0,15)#12m has higher variacne then ldlcm due to possible life style change after entering the trial
  data0$ldlc12<-data0$ldlc12m+rnorm(n,0,16)#add vairance of rgtm
  #treatment arm
  for(i in 1:100){
    #print(i)
    if(n1<n){
      ldlcm=rnorm(n = n*osp,mean = 116,sd = 14.2)
      d1<-data.frame(ldlcm,ldlcb=rnorm(n = n*osp,mean = ldlcm,sd = 16))
      d1<-d1%>%filter(ldlcb<cutoff)
      data1<-rbind(data1,d1)
      n1<-nrow(data1)
    } else {
      break
    }
  }
  data1<-data1[1:n,]
  data1$ldlc12m<-data1$ldlcm - 0.7*data1$ldlcb +30 + rnorm(n,0,15)
  data1$ldlc12<-data1$ldlc12m+rnorm(n,0,sd = 16)
  data0$drug = 0
  data1$drug = 1
  data = rbind(data0,data1)
  #data<-data%>%mutate(ldlc_change = ldlc12 - ldlcb,ldlc_change.m = ldlc12m - ldlcb)
  return(data)
}


```

## Parameters
```{r}

#Set Up
p<<-1.4
#outcome regression AFT
theta0<<-7.5
theta1<<-0.12
theta2<<-0.015#should be negative
theta3<<--0.023
theta4<<--0.02

theta0_ni<<-7.7
theta1_ni<<-0.59
theta2_ni<<-0.0032 #should be negative
cutoff<-c(90,110,130,150,1000)
```


## Identification of regression towards the mean (RGTM)

1. Measurement change with certain pattern, to the mean(observed)\n

2. Condition on a certain group, variance increase with time\n

3. Different from selection bias, as selection bias focuses on mean\n

## Properities of RGTM (What changes it)

In simple case, assume different measurement are normally distributed. If more data is available, more complex model could be used to model individual variation. If measurment are normally distributed, three aspect could affect RGTM \n
 
1. Mean of the individual.\n

2. Individual  How the biomarkers of individual change over time. \n

3. Selection. How the data is selected. In JUPITER Trial, the selection criteria is LDLC below 130. Sometimes refered as cutoff point. \n

## Simulation 


### Example

Apply no selection \n
```{r}
#test case, no cutoff
data<-simulation.function.v42(20000,cutoff = 1000)
data$ldlc_change<-data$ldlc12 - data$ldlcb
data$ldlc_change_m<-data$ldlc12m - data$ldlcm

data0<-data%>%filter(drug == 0)
data1<-data%>%filter(drug == 1)
hist(data0$ldlc_change)
plot(data0$ldlcb,data0$ldlc_change)
```

Select 90 and below \n
```{r}
data<-simulation.function.v42(20000,cutoff = 90)
data$ldlc_change<-data$ldlc12 - data$ldlcb
data$ldlc_change_m<-data$ldlc12m - data$ldlcm

data0<-data%>%filter(drug == 0)
data1<-data%>%filter(drug == 1)
hist(data0$ldlc_change)

plot(data0$ldlcb,data0$ldlc_change)
```

Apply a selection results a different behavior on LDLC 




```{r}



fit<-lm(data,formula = ldlc_change_m~ldlc_change+ldlcb+drug)
plot(fit$residuals)

data$ldlc_change_m_pred<-predict(fit,data)
plot(data$ldlc_change_m,data$ldlc_change_m_pred)#mean prediction works

```

